---
title: "Reading Japan Post's KEN_ALL, legacy and modern"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Japan Post publishes every postal code in the country as a CSV, and has done
for decades. It comes in two editions: an original in Shift-JIS, and a newer
one in UTF-8. Between them they exercise most of what makes real CSV files
harder than the examples in a parser's test suite, so they make a good place
to show what `zucsv` does and — just as usefully — where it deliberately
disagrees with the other readers.

```{r download, include = FALSE}
dir <- file.path(tempdir(), "kenall")
dir.create(dir, showWarnings = FALSE)

fetch <- function(url, dest) {
  if (file.exists(dest)) return(TRUE)
  ok <- tryCatch({
    utils::download.file(url, dest, quiet = TRUE, mode = "wb")
    file.exists(dest) && file.size(dest) > 1e6
  }, error = function(e) FALSE, warning = function(w) FALSE)
  ok
}

# Japan Post's own CDN has been returning a cached 404 for its whole
# download tree, so these come from the Internet Archive's copies of the
# official zips.
base <- "https://web.archive.org/web/2025id_/https://www.post.japanpost.jp/zipcode/dl"
sjis_zip <- file.path(dir, "ken_all.zip")
utf8_zip <- file.path(dir, "utf_ken_all.zip")
have <- fetch(paste0(base, "/kogaki/zip/ken_all.zip"), sjis_zip) &&
        fetch(paste0(base, "/utf/zip/utf_ken_all.zip"), utf8_zip)

if (have) {
  utils::unzip(sjis_zip, exdir = dir)
  utils::unzip(utf8_zip, exdir = dir)
  sjis <- file.path(dir, "KEN_ALL.CSV")
  utf8 <- file.path(dir, "utf_ken_all.csv")
  have <- file.exists(sjis) && file.exists(utf8)
}
```

```{r no-data, echo = FALSE, results = "asis", eval = !have}
cat("> **The data could not be downloaded**, so the results below are not",
    "shown. Everything in the prose still applies; re-run the article with",
    "network access to reproduce it.\n")
```

## The legacy file is not UTF-8, and that matters

`KEN_ALL.CSV` is Shift-JIS, has no header, and uses CRLF. Ask `zucsv` to read
it and you get an error rather than a data frame:

```{r sjis-zucsv, eval = have, error = TRUE}
library(zucsv)
read_csv(sjis, header = FALSE)
```

That is `zucsv` doing what it is supposed to do. Row 1, column 4 is the first
field in the file containing a byte that is not valid UTF-8 — the half-width
katakana `ﾎｯｶｲﾄﾞｳ`, at byte offset 26. The message names the record and the
column, so you know both *what* is wrong and *where*.

It is worth comparing, because the alternatives are not obviously worse until
you look closely.

```{r sjis-others, eval = have}
first_bytes <- function(x) paste(utils::head(charToRaw(x), 6), collapse = " ")

# data.table and vroom both accept the file and return something
dt <- data.table::fread(sjis, header = FALSE, showProgress = FALSE)
vr <- vroom::vroom(sjis, delim = ",", col_names = FALSE,
                   show_col_types = FALSE, altrep = FALSE)

c(fread = first_bytes(dt[[7]][1]),
  vroom = first_bytes(vr[[7]][1]),
  truth = first_bytes(
    utils::read.csv(sjis, header = FALSE, fileEncoding = "CP932")[1, 7]))
```

Column 7 of the first record is the prefecture name, 北海道. Read correctly it
is nine bytes of UTF-8. `fread` and `vroom` hand back the raw Shift-JIS bytes
with no encoding mark, so nothing has failed yet — the failure arrives later,
somewhere else, in whatever code first tries to treat those bytes as text.

`utils::read.csv()` does error, but it takes about 100× longer to do it, and
the message points at a byte sequence rather than a location. It is also an
error you cannot print safely: the message itself contains the invalid bytes.

### Reading it properly

Convert the encoding first, then read:

```{r sjis-converted, eval = have}
converted <- file.path(dir, "KEN_ALL-utf8.csv")
writeLines(
  iconv(readLines(sjis, encoding = "CP932", warn = FALSE), "CP932", "UTF-8"),
  converted, useBytes = TRUE
)

old <- read_csv(converted, header = FALSE)
dim(old)
old[1, c(3, 7, 8, 9)]
```

## The modern file

`utf_ken_all.csv` is the same data, normalised to UTF-8 and to one row per
record:

```{r utf8, eval = have}
new <- read_csv(utf8, header = FALSE)
dim(new)
vapply(new, function(x) class(x)[1], "")
```

Kanji comes back marked as UTF-8, with the character counts you would expect:

```{r utf8-text, eval = have}
x <- new[[7]][1]
c(value = x, encoding = Encoding(x), nchar = nchar(x), bytes = nchar(x, "bytes"))
```

## The postal-code trap

Column 3 is the seven-digit postal code, and it is the reason this file is a
good teacher. Hokkaido's first entry is `0600000`:

```{r zeros, eval = have}
as_text <- read_csv(utf8, header = FALSE, col_types = "character")
as_text[[3]][1]

# inferred, it is a number
new[[3]][1]

# how many codes would lose a leading zero
mean(substr(as_text[[3]], 1, 1) == "0")
```

About one code in eight begins with a zero, and reading the column as a number
destroys it. `zucsv` infers `integer` here, and so do `read.csv()` and `fread`
— this is not a difference between readers, it is what "looks like a number"
means. The fix is to say so:

```{r zeros-fixed, eval = have}
fixed <- read_csv(utf8, header = FALSE,
                  col_types = c("integer", rep("character", 8),
                                rep("integer", 6)))
fixed[[3]][1]
```

Leading zeros being integer syntax is a deliberate choice in `zucsv`, matching
`utils::type.convert()`. `col_types` is the documented escape, and this file is
exactly the case it exists for.

## Whitespace is never trimmed

Column 2 is the old five-digit code, space-padded to five characters:

```{r padded, eval = have}
as_text[[2]][1]
class(new[[2]])
class(utils::read.csv(utf8, header = FALSE)[[2]])
```

`zucsv` leaves `"060  "` as text, because it is text — nothing is trimmed
before deciding a type. Base R trims first and calls it the number 60. Both are
defensible; `zucsv` prefers not to silently discard bytes that were in the
file.

## One more thing the two editions differ on

The legacy file splits long town names across consecutive records that share a
postal code; the modern one keeps each record whole.

```{r splitting, eval = have}
c(old_rows = nrow(old), new_rows = nrow(new))

# the longest town field in the modern file
i <- which.max(nchar(as_text[[9]]))
nchar(as_text[[9]][i])
```

That is a property of the data, not of CSV syntax — every reader sees the same
row counts, and reassembling split records is the caller's job either way.

## What to take from this

* **Encoding problems should be loud.** `zucsv` refuses invalid UTF-8 and says
  where it is. Deferring the failure to whatever code touches the string next
  is not a kindness.
* **`col_types` is not a performance knob.** On this file it is the difference
  between keeping postal codes and quietly corrupting one in eight of them.
* **Type inference cannot read your mind.** "Looks like a number" is a
  syntactic question, and every reader here answers it the same way.
