---
title: "Reading CSV files with zucsv"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(zucsv)
```

`zucsv` reads delimited text files into ordinary R data frames, using the
[`zsv`](https://github.com/liquidaty/zsv) C parser as its backend. This article
covers the whole package — there are two exported functions — and, more usefully,
the handful of decisions that make it behave differently from the readers you
already have.

## Why another CSV package

R is not short of CSV readers. `zucsv` is not trying to replace `data.table` or
`readr`; it is a smaller thing with different priorities:

- **Semantics you can check.** Numeric conversion is bit-identical to
  `as.numeric()`, by construction rather than by coincidence.
- **Errors where the problem is.** Invalid UTF-8, an embedded NUL or a ragged
  row is an error naming the record and column, not a value that fails later
  somewhere else.
- **No silent repair.** Duplicate column names stay duplicated, nothing is
  whitespace-trimmed, and a value that will not convert to a type you asked for
  is an error rather than `NA`.
- **No dependencies, no system library.** The parser is bundled and pinned.
- **A small surface.** One function to read a file and one to report how it
  would be read, five arguments between them, and a documented list of what
  `zucsv` deliberately does not do.

If you want the fastest possible reader, use `data.table::fread()`. If you want
tidyverse integration, use `readr`. `zucsv` optimises for predictability.

## The API

```{r, eval = FALSE}
read_csv(file, header = TRUE, delimiter = ",", na = c("", "NA"), col_types = NULL)
```

| argument | what it does |
|---|---|
| `file` | path to a local file; `~` is expanded |
| `header` | `TRUE` first record supplies names, `FALSE` generate `V1`…`Vn`, `NA` detect which |
| `delimiter` | any single ASCII byte — comma, tab, semicolon, pipe |
| `na` | exact cell values to read as missing, or `NULL` for none |
| `col_types` | force `logical`, `integer`, `double` or `character` |

`header = NA` is for files whose shape you do not know in advance. It reads the
first record as names unless one of its unquoted fields is a number or a
logical — so `id,name,score` is a header and `1,Ada,9.5` is data. It judges
that one record and nothing else, which is cheap and predictable but less
informed than the equivalent in `data.table::fread()` or DuckDB; `?read_csv`
lists the three shapes where the verdicts differ. When you know the answer,
say so.

```{r}
path <- tempfile(fileext = ".csv")
writeLines(c(
  "id,name,score,active",
  "1,Ada,9.5,TRUE",
  "2,Linus,8.0,FALSE",
  "3,Grace,,TRUE"
), path)

df <- read_csv(path)
df
str(df)
```

If you want to see those decisions before committing to them, `sniff_csv()`
reports them without building any columns --- and its output feeds straight back
in, which turns an inferred read into a strict one:

```{r}
s <- sniff_csv(path)
str(s)

# same data frame, but nothing is inferred this time
identical(read_csv(path, header = s$header, col_types = s$col_types),
          read_csv(path))
```

That second form is worth reaching for in a script that runs unattended: pin the
schema once, and a file that changes shape becomes an error instead of a
differently-typed column.

Other delimiters need no separate function:

```{r}
tsv <- tempfile(fileext = ".tsv")
writeLines(c("a\tb", "1\t2"), tsv)
read_csv(tsv, delimiter = "\t")
```

## Type inference

Columns are inferred over `logical`, `integer`, `double` and `character`, in
that order, using **every** value in the column rather than a sample. A column
is only narrower than `character` if every non-missing value in the whole file
fits.

```{r}
types <- tempfile(fileext = ".csv")
writeLines(c("lgl,int,dbl,chr", "TRUE,1,1.5,x", "FALSE,2,2.5,y"), types)
vapply(read_csv(types), typeof, "")
```

Only four spellings are logical — `TRUE`, `FALSE`, `true`, `false`. `T`, `yes`
and `1` are not, because in real data they are far more often something else.

```{r}
spellings <- tempfile(fileext = ".csv")
writeLines(c("a,b,c", "T,yes,TRUE", "F,no,FALSE"), spellings)
vapply(read_csv(spellings), typeof, "")
```

### Forcing types

`col_types` takes one value for every column, or one value for all of them. A
cell that will not convert is an error naming the row, the column and the
value:

```{r, error = TRUE}
read_csv(path, col_types = c("integer", "character", "double", "logical"))

read_csv(path, col_types = "integer")
```

That strictness is the point. A reader that turned `"Ada"` into `NA` here would
be losing data quietly.

## Four things that surprise people

These are all deliberate, and all of them are the kind of thing worth knowing
before you meet them in a 200-megabyte file.

**Blank lines are skipped, everywhere.** Leading, trailing or in the middle —
they are not rows.

```{r}
blanks <- tempfile(fileext = ".csv")
writeLines(c("a,b", "1,2", "", "3,4", ""), blanks)
nrow(read_csv(blanks))
```

A line of only whitespace or only delimiters is *not* blank: it is a record of
empty cells, and must match the table's width. In a one-column file, write an
intentional empty cell as `""` — a bare empty line is skipped.

**Nothing is trimmed.** `" 42"` is text, because it has a space in it.

```{r}
padded <- tempfile(fileext = ".csv")
writeLines(c("a", " 42", "1"), padded)
read_csv(padded)$a
```

**Leading zeros are integers.** This matches `utils::type.convert()`, and it is
how postal codes and zip codes lose their leading zero in every reader. Use
`col_types` when the column is an identifier rather than a number.

```{r}
codes <- tempfile(fileext = ".csv")
writeLines(c("zip", "0600000", "1000001"), codes)
read_csv(codes)$zip
read_csv(codes, col_types = "character")$zip
```

**Column names are kept exactly**, duplicates and empty strings included. No
name repair.

```{r}
dupes <- tempfile(fileext = ".csv")
writeLines(c("a,a,", "1,2,3"), dupes)
names(read_csv(dupes))
```

## Errors say where

Structural problems name the record and the column. Records are numbered as
they appear, header included, so the numbers match what you see in an editor.

```{r, error = TRUE}
ragged <- tempfile(fileext = ".csv")
writeLines(c("a,b", "1,2", "3,4,5"), ragged)
read_csv(ragged)
```

Text that cannot be carried into an R string is an error too, rather than
something that fails later:

```{r, error = TRUE}
bad <- tempfile(fileext = ".csv")
con <- file(bad, "wb")
writeBin(c(charToRaw("a,b\n"), as.raw(c(0xFF, 0xFE)), charToRaw(",2\n")), con)
close(con)
read_csv(bad)
```

The [KEN_ALL article](japanese-postal-codes.html) shows why that matters on a
real file, and what the other readers do with the same input.

## Missing values

`na` is a set of exact cell values, matched after unquoting. Quoted and
unquoted forms are therefore the same thing, and nothing is trimmed first.

```{r}
miss <- tempfile(fileext = ".csv")
writeLines(c("a,b,c", 'NA,"NA", NA'), miss)
read_csv(miss)

read_csv(miss, na = NULL)
```

## Performance

Faster than `utils::read.csv()` by 2–12× depending on shape; slower than
`data.table::fread()`, which is multi-threaded and single-pass. The
[benchmark results](https://github.com/pedrobtz/zucsv/blob/main/bench/RESULTS.md)
have the numbers and the method, including why the comparison against `vroom`
needs care.

## What it does not do

v0.1 reads local files and nothing else. There is no `write_csv()`, no URL or
connection input, no compressed files, no date parsing, no column selection, no
`skip` or `n_max`, and no `encoding` argument — convert the file first, as the
KEN_ALL article shows. The full list, and the reasoning, is in the package's
design document.
